Exploring Determinants of Child Mortality Globally (2022): The Role of Vaccination, Healthcare Spending, and Nutrition

### 2025-04-14 paper/paper.qmd
Method

# Method

@@

Here we will investigate two loss functions, ranknet loss and focal loss. Ranknet loss is sepcified as follow: Given training data $x_1$, $x_2$, .... , $x_n$, and their rankings, $y_1$, $y_2$, ..., $y_n$, we will investigate all combinations of ordered pairs ($x_i$, $x_j$). The total loss is: 
$$
L = \sum_{i < j} - y_{ij} \log P(i > j) - (1 - y_{ij}) \log (1 - P(i > j))
$$
Where $y_ij$ is 1 if $x_i$ is ranked higher than $y_j$, 0 if $x_i$ is ranked lower than $y_j$ (here we will assume no $x_i$ can have the same ranking as $x_j$ if i is not equal to j). $ P(i > j) = \frac{1}{1 + e^{-(s(x_i) - s(x_j))}} $ is the probability that $ x_i $ is ranked higher than $ x_j $. s() is the rankning score assigned to i, higher s higher ranking.

>>>>

# Method {#sec-method}

@@

Here we will investigate two loss functions, ranknet loss and focal loss. Ranknet loss is sepcified as follow: Given training data $x_1$, $x_2$, .... , $x_n$, and their rankings, $y_1$, $y_2$, ..., $y_n$, we will investigate all combinations of ordered pairs ($x_i$, $x_j$). The total loss is: 
$$
L = \sum_{i < j} - y_{ij} \log P(i > j) - (1 - y_{ij}) \log (1 - P(i > j))
$$
Where $y_ij$ is 1 if $x_i$ is ranked higher than $y_j$, 0 if $x_i$ is ranked lower than $y_j$ (here we will assume no $x_i$ can have the same ranking as $x_j$ if i is not equal to j). $P(i > j) = \frac{1}{1 + e^{-(s(x_i) - s(x_j))}}$ is the probability that $x_i$ is ranked higher than $x_j$. s() is the rankning score assigned to i, higher ranking cores means higher ranking. This loss aims to peanalize the pairs of data $x_i$ and $x_j$ where $x_i$ should be ranked higher than $x_j$ but instead $x_i$ is ranked lower by the model for $x_j$.
